---
title: "neighborhood selection"
author: judy cheng
date: april 7,2018
output: html_notebook
---
##component study:neighborhood selection
#what is it:
determine which other users' data to be used in the prediction for a user
#motivation: 
1.improve accuracy--high correlates can be exceptionally more valuable as predictors.
2.improve computational time--commercial collaborative filtering are beginning to handle millions of users, making consideration of every user impossible.
#two methods:
1)
correlation-thresholding:
definition:set an absolute correlation threshold, where all neighbors with correlations above the given threshold are selected.
advantages:a high threshold gives you high correlates
disadvantages:a high threshold gives you a small neighborhood that limits prediction coverage for many items(because for many users high correlates are not available)
2)top-n-neighbors: 
definition:pick the best n correlates for a given n.
advantages:does not limit prediction coverage as correlation-thresholding does.
disadvantages:a high n will result in noise(many low correlates) for users who have high correlates; picking a low n will result in poor prediction for users who do not have many high correlates.
#our results:
for correlation-thresholding, we tested absolute correlations above 0.2,0.4,0.6,0.8
for top-n-neighbors, we tested 10%, 30%, 50%, and 70% of users
#future possible considerations:
we could combine the two methods, i.e. combining a low threshold with a high n or a high threshold with a low n.
```{r}
#set working directory and load functions
setwd("/Users/Judy/Documents/GitHub/project-3-algorithms-project-4-algorithms-group-4/data")
source("../lib/old_functions.R")

########MS data#######


#load data
load("../data/MS_UI.RData")
load("../data/MS_sim.RData")

#make some vectors for computational times
comp_time<-c() #MS topn
MS_threshold_time<-c() #MS threshold

#do the initial matrix preparation common to both methods of neighborhood selection
#pred_initial(MS_UI)


####method 1:top n neighbors####
#total number of neighbors
all_neighb<-nrow(MS_UI)
#several values of n to try out--10%,30%,50%,and 70% of all neighbors
n<-c(0.1,0.3,0.5,0.7)

#get the prediction matrices and computational time for different values of n
comp_time<-c(comp_time, system.time(MS_topn_10percent <- pred_topn(MS_UI, MS_sim, trunc(all_neighb*n[1]))))
View(MS_topn_10percent)
save(MS_topn_10percent, file = "MS_topn_10percent.RData")

comp_time<-c(comp_time, system.time(MS_topn_30percent <- pred_topn(MS_UI, MS_sim, trunc(all_neighb*n[2]))))
save(MS_topn_30percent, file = "MS_topn_30percent.RData")

comp_time<-c(comp_time, system.time(MS_topn_50percent <- pred_topn(MS_UI, MS_sim, trunc(all_neighb*n[3]))))
save(MS_topn_50percent, file = "MS_topn_50percent.RData")

comp_time<-c(comp_time, system.time(MS_topn_70percent <- pred_topn(MS_UI, MS_sim, trunc(all_neighb*n[4]))))
save(MS_topn_70percent, file = "MS_topn_70percent.RData")



####method 2: correlation thresholding####
#several thresholds to try out
th<-c(0.2,0.4,0.6,0.8)

#get the prediction matrices and computational time for different thresholds
MS_threshold_time<-c(MS_threshold_time, system.time(MS_thr_0.2 <- pred_threshold(MS_UI, MS_sim, th[1])))
save(MS_thr_0.2, file = "MS_thr_0.2.RData")

MS_threshold_time<-c(MS_threshold_time, system.time(MS_thr_0.4 <- pred_threshold(MS_UI, MS_sim, th[2])))
save(MS_thr_0.4, file = "MS_thr_0.4.RData")

MS_threshold_time<-c(MS_threshold_time, system.time(MS_thr_0.6 <- pred_threshold(MS_UI, MS_sim, th[3])))
save(MS_thr_0.6, file = "MS_thr_0.6.RData")

MS_threshold_time<-c(MS_threshold_time, system.time(MS_thr_0.8 <- pred_threshold(MS_UI, MS_sim, th[4])))
save(MS_thr_0.8, file = "MS_thr_0.8.RData")


```

```{r}
########movie data##########
#load data
load("/Users/Judy/Documents/GitHub/project-3-algorithms-project-4-algorithms-group-4/data/movie_UI.RData")
load("/Users/Judy/Documents/GitHub/project-3-algorithms-project-4-algorithms-group-4/data/movie_sim.RData")

#make some vectors to record computational time
movie_topn_time<-c()
movie_threshold_time<-c()


#do the initial matrix preparation common to both methods of neighborhood selection
#pred_initial(movie_UI)

####method 1:top n neighbors####
#total number of neighbors
all_neighb<-nrow(movie_UI)
#several values of n to try out--10%,30%,50%,and 70% of all neighbors
n<-c(0.1,0.3,0.5,0.7)

#get the prediction matrices and computational time for different values of n
movie_topn_time<-c(movie_topn_time, system.time(movie_topn_10percent <- pred_topn(movie_UI, movie_sim, trunc(all_neighb*n[1]))))
save(movie_topn_10percent, file = "movie_topn_10percent.RData")

movie_topn_time<-c(movie_topn_time, system.time(movie_topn_30percent <- pred_topn(movie_UI, movie_sim, trunc(all_neighb*n[2]))))
save(movie_topn_30percent, file = "movie_topn_30percent.RData")

movie_topn_time<-c(movie_topn_time, system.time(movie_topn_50percent <- pred_topn(movie_UI, movie_sim, trunc(all_neighb*n[3]))))
save(movie_topn_50percent, file = "movie_topn_50percent.RData")

movie_topn_time<-c(movie_topn_time, system.time(movie_topn_70percent <- pred_topn(movie_UI, movie_sim, trunc(all_neighb*n[4]))))
save(movie_topn_70percent, file = "movie_topn_70percent.RData")



####method 2: correlation thresholding####
#several thresholds to try out
th<-c(0.2,0.4,0.6,0.8)

#get the prediction matrices and computational time for different thresholds
movie_threshold_time<-c(movie_threshold_time, system.time(movie_thr_0.2 <- pred_threshold(movie_UI, movie_sim, th[1])))
save(movie_thr_0.2, file = "movie_thr_0.2.RData")

movie_threshold_time<-c(movie_threshold_time, system.time(movie_thr_0.4 <- pred_threshold(movie_UI, movie_sim, th[2])))
save(movie_thr_0.4, file = "movie_thr_0.4.RData")


movie_threshold_time<-c(movie_threshold_time, system.time(movie_thr_0.6 <- pred_threshold(movie_UI, movie_sim, th[3])))
save(movie_thr_0.6, file = "movie_thr_0.6.RData")


movie_threshold_time<-c(movie_threshold_time, system.time(movie_thr_0.8 <- pred_threshold(movie_UI, movie_sim, th[4])))
save(movie_thr_0.8, file = "movie_thr_0.8.RData")

```
