---
title: "Collaborative Filtering"
author: "Group 4"
date: "4/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Step 0: Load Data
```{r}
# setwd("/Users/Nicole/Documents/GitHub/project-3-algorithms-project-4-algorithms-group-4/doc")
# Change working directory 

load(file = "../data/MS_UI.RData")
load(file = "../data/movie_UI.RData")
```


## Task: Memory Based Collaborative Filtering

### Comparing different similarity weights
#### Pearson correlation
```{r}
# Weights
load(file = "../data/MS_sim.RData")
load(file = "../data/movie_sim.RData")

# Predictions
load(file ="../data/MS_pred.RData")
load(file = "../data/movie_pred.RData")
```

#### Spearman's correlation
```{r}
MS_train <- MS_UI
Movie_train <- movie_UI 


```

#### Vector similarity
```{r}

```


#### Entropy
```{r}

```


#### Mean-Square Difference
```{r}

```


#### SimRank
```{r}

```


## Task: Group 4 Specific Assignment - Neighbors 
```{r}

```



## Task: Model Based Collaborative Filtering

### Step 1: Build EM algorithm 
```{r}
## Initialize parameters


## E-step


## M-step
```


### Step 2: Cross validation to choose the best number of clusters
```{r}

```


### Step 3: Predict ratings
```{r}

```


### Step 4: Calculate test error
```{r}

```

