---
title: "Collaborative Filtering"
author: "Nicole Smith"
date: "4/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/Users/Nicole/Documents/GitHub/project-3-algorithms-project-4-algorithms-group-4/doc")

load(file = "../data/MS_UI.RData")
load(file = "../data/movie_UI.RData")
```

```{r Matrix Calculations}
visit_nums <- rowSums(MS_UI, na.rm = TRUE)

# table(visit_nums)
# mean(visit_nums)
# median(visit_nums)

total_ratings <- rowSums(movie_UI, na.rm = TRUE)

# table(total_ratings)
# mean(total_ratings)
# median(total_ratings)
```


```{r Simularity Weights of the Users}
# Pearson correlations
load(file = "../data/MS_sim.RData")
load(file = "../data/movie_sim.RData")

# Spearman correlations

# Vector similarity (Cosine)

# Entropy-based

# Mean-square difference

# SimRank

```

```{r Calculating Predictions for Users}
# From Pearson Simularity Weights
load(file ="../data/MS_pred.RData")
load(file = "../data/movie_pred.RData")
```

Group 4:
- After calculating similarity weights, we select which other users' data are used in computing the predictions (currently, all of them).
- There is evidence that selecting a subset of users improves accuracy. 
- Moreover, when there are millions of users, using them all is infeasible. 

Explore: Will prediction accuracy improve if we select the best neighbors of the active user to use in calculating predictions?


```{r Model-Based Algorithm}
library(EMCluster)
```

